---
title: "quarto_homework_3"
format: html
editor: visual
author: Marcelo Miller
date: 5/21/25
---

```{r}
library(tidyverse)
library(flextable)
library(gt)
library(janitor)
library(MuMIn)
library(scales)
library(here)
library(dplyr)
library(readr)
```

<https://github.com/marcelotmiller/ENVS_193DS_homework-03.git>

#1

a.)

The data that I have gathered for my caffeine study consist primarily of binomial categorical data (hw/exam due and exercised) and continuous variables (caffeine amount), along with ordinal variables (date and time). The primary way that I would be comparing a response and independent variables are by comparing the means of caffeine amounts on days that I have homework/exam due vs those that I don't, with the intention of proving or denying the suspicion I have that I consume more caffeine on academically demanding days.

b.)

```{r}

library(tidyverse)

# load and clean the data
data <- read_csv("data_trial_june1st")

clean_data <- data |>
  filter(!is.na(`Caffeine (mg)`), !is.na(`HW/Exam Due`)) |>
  select(`HW/Exam Due`, `Caffeine (mg)`)

# summarize the mean of the caffeine consumption by Yes/No groups
caffeine_summary <- clean_data |>
  group_by(`HW/Exam Due`) |>
  summarize(mean_caffeine = mean(`Caffeine (mg)`), .groups = "drop")

# create plot with two categories and clear ticks for y-axis
ggplot(clean_data, aes(x = `HW/Exam Due`, y = `Caffeine (mg)`, color = `HW/Exam Due`)) +
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +
  stat_summary(fun = mean, geom = "crossbar", width = 0.5, fatten = 2, color = "black") +
  scale_color_manual(values = c("YES" = "chartreuse3", "NO" = "brown1")) +
  scale_y_continuous(limits = c(0, 400), breaks = seq(0, 400, 50)) +
  labs(
    title = "Caffeine Consumption on Days With and Without Homework or Exams",
    x = "Homework or Exam Due?",
    y = "Caffeine Intake (mg)", 
    caption = "Figure 1. Caffeine intake on days that involve an exam or homework being due and days that do not. \nEach point represents the caffeine consumption for one day (mg). The horizontal black bars represent the means for each category. "
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", hjust = 0.5, size = 12), plot.caption = element_text(size = 10, hjust = 0, margin = margin(t = 10))
  )
```
